---
title: "Day_1"
output: html_document
date: "2025-06-03"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(devtools)
library(usethis)

setwd("C:/Users/swulfing/Documents/GitHub/UMassD/Coursework_RPackages")

has_devel() # This will tell you if you have the infrastructure to make a package (rtools properly instaled)

# Want 3.1.0 version of usethis

usethis::git_sitrep() # info about github, checks that you're set up
# usethis::use_git_config(user.name = "blah blah", user.email = "blah")
usethis::git_vaccinate() # Make sure Vaccinate = TRUE
usethis::create_github_token()
gitcreds::gitcreds_set() #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
```

# TO DO

LINK YOUR GIT: https://happygitwithr.com/rstudio-see-git

# Links:

 - https://forwards.github.io/package-dev/workshops/summer-2025-cohort-2.html

 - https://r-pkgs.org/
 
 # Notes:
 
 Today:
 
  - Why write packages
  - Package structure and states
  - Package development setup
  - Creating a package with a working function
  
  
Scropt vs package

Script:

 - one-off analysis
 - .R extension
 - Documentatnion in #
 - Runes lines or source files
 
Package:
 - reusable components
 - defined by presence of description file
 - imports defined by namespace file
 - documentation in file and Roxygen comments
 - install and restart
 
 Names of files and directories MUST follow specification in the Writing R Exensions manual
 All have: DESCRIPTIOM, NAMESPACE, R, mypackage.rproj
 
 Package states
  - source code
  - bundled - compressed file. Don't usually have to make them, created by CMD build
  - Binary - Compressed copy of package once installed. Single file. Platfor specific (zip for windows). Developers submit a bundle to CRAN; CRAN makes and distributes binaries
  - Installed - a directory of files in a lib, any C/ C++/fortran in compled form. Help files, code, optional data. install.packages() can install from source or from binary
  - in-memory - if installed, library() makes the function available to memory
  

```{r buildPackage}
usethis::create_package("C:/Users/swulfing/Documents/GitHub/UMassD/Coursework_RPackages/animalsounds")
```




















