---
title: "WHAMMSETest"
output: html_document
date: "2025-05-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

write.dir <- "C:/Users/swulfing/OneDrive - University of Massachusetts Dartmouth/Documents/GitHub/UMassD/GavinMSECourse"
setwd(write.dir)

library(tidyverse)
library(ggdist)
library(Hmisc)
library(mvtnorm)

```
# Changes from code:

1. Observe function (in page 13 of my_first_mse PDF or code chunk labeled "applyMSE") needs to change from $-0.5*sigma^2$ to $-0.5*sigma**2$

2. Overshooting the TAC in page 19 of PDF or eval function on "applyMSE" (fxn is also called in "project_fix10percent" and "MSE_altHCRs") I couldn't figure out how to use the ... as reading in extra vars so I added overshoot to the functions and function calls

```{r OM_devel}

data.years <- 1991:2013
harvest <- c(0.1,3,15,52,76,139,95,93,84,93,86,103,104,
             92,46,67,59,30,54,59,47,33,44)
index <- c(NA,NA,NA,NA,NA,NA,NA,NA,935,NA,1057,NA,678,NA,
           420,NA,554,NA,458,NA,474,NA,280)

plot(data.years,index, pch=19,xlab="Year",ylab="Million tonnes (B/C)",
     ylim=c(0,1200))
lines(data.years,harvest,lty=2,lwd=2)
```

```{r fxns}
# Logistic Production FXN
schaefer <- function(B,C,K,r) {
  #function schaefer takes the current biomass, a catch,
  #and the model parameters to compute next year's biomass
  res <- B + B * r * (1 - B/K) - C
  return(max(0.001,res)) # we add a constraint to prevent negative biomass
}

# Biomass projection
dynamics <- function(pars,C,yrs) {
  # dynamics takes the model parameters, the time series of catch,
  # & the yrs to do the projection over
  # first extract the parameters from the pars vector (we estimate K in log-space)
  K <- exp(pars[1])
  r <- exp(pars[2])
  # find the total number of years
  nyr <- length(C) + 1
  # if the vector of years was not supplied we create
  # a default to stop the program crashing
  if (missing(yrs)) yrs <- 1:nyr
  #set up the biomass vector
  B <- numeric(nyr)
  #intialize biomass at carrying capacity
  B[1] <- K
  # project the model forward using the schaefer model
  for (y in 2:nyr) {
    B[y] <- schaefer(B[y-1],C[y-1],K,r)
  }
  #return the time series of biomass
  return(B[yrs])
  #end function dynamics
}

# We want to codition OM by estimate params based on historical biomass index data
# Make a fxn that shows how well current params fit data
# function to calculate the negative log-likelihood
nll <- function(pars,C,U) { #this function takes the parameters, the catches, and the index data
  sigma <- exp(pars[3]) # additional parameter, the standard deviation of the observation error
  B <- dynamics(pars,C) #run the biomass dynamics for this set of parameters
  Uhat <- B #calculate the predicted biomass index - here we assume an unbiased absolute biomass estimate
  output <- -sum(dnorm(log(U),log(Uhat),sigma,log=TRUE),na.rm=TRUE) #calculate the negative log-likelihood
  return(output)
  #end function nll
}

# FXN to perform assessment and estimate OM params
assess <- function(catch,index,calc.vcov=FALSE,pars.init) {
  # assess takes catch and index data, initial values for the parameters,
  # and a flag saying whether to compute uncertainty estimates for the model parameters
  #fit model
  # optim runs the function nll() repeatedly with differnt values for the parameters,
  # to find the values that give the best fit to the index data
  res <- optim(pars.init,nll,C=catch,U=index,hessian=TRUE)
  # store the output from the model fit
  output <- list()
  output$pars <- res$par
  output$biomass <- dynamics(res$par,catch)
  output$convergence <- res$convergence
  output$nll <- res$value
  if (calc.vcov)
    output$vcov <- solve(res$hessian)
  return(output)
  #end function assess
}

```

```{r run_assess}
ini.parms <- c(log(1200), log(0.1), log(0.3))

redfish <- assess(harvest,index,calc.vcov=TRUE,ini.parms)
redfish

# Max likelihood and param estimates
biomass.mle <- redfish$biomass
print(biomass.mle)

pars.mle <- redfish$pars
print(exp(pars.mle))

# Obtain alt params of OM. Use statistical uncertainty from estimation by sampling param sets from estimated variance-covariance mtx

#define the number of iterations for the MSE
niter <- 200
#set up a storage matrix for our alternative parameter sets
pars.iter <- matrix(NA,nrow = niter, ncol=3)
colnames(pars.iter) <- c("log_K","log_r","log_sigma")
# generate the sets of parameter values
for (i in 1:niter) {
  pars.iter[i,] <- mvtnorm::rmvnorm(1, mean = redfish$pars,
                                    sigma = redfish$vcov)
}
# Now generate replicate model outputs
biomass.iter <- data.frame()
for (i in 1:niter) {
  #here we calculate the biomass trajectory for each of the above sampled parameter vectors
  biomass.iter <- rbind(biomass.iter,
                        data.frame(year = seq(min(data.years),
                                              max(data.years)+1),
                                   biomass = dynamics(pars.iter[i,], harvest),
                                   iter = i))
}
biomass.iter

```

```{r plot}

biomass.iter %>%
  group_by(year) %>%
  median_qi(biomass, .width = c(.5, .8, .95)) %>%
  ggplot() +
  geom_lineribbon(aes(x = year, y = biomass, ymin = .lower, ymax = .upper),
                  show.legend = FALSE) +
  scale_fill_brewer() +
  #theme_bw() +
  geom_line(aes(y=harvest,x=year), data = tibble(harvest = harvest,
                                                 year = data.years),lty=2) +
  geom_point(aes(y=index, x=year), data = data.frame(index=index,
                                                     year = data.years)) +
  #geom_line(aes(y=biomass,x=year,group=iter,col=iter),data = subset(biomass.iter,iter%in%1:10)) +
  ylab("Estimated B and C (million tonnes)") +
  theme_bw() +
  guides(scale = "none")

```

```{r applyMSE}
proj.years <- 2014:2034

##### Data generation
observe <- function(biomass, sigma) {
  biomass * exp(rnorm(1, -0.5*sigma**2, sigma))
}

# HCR
control.pars <- list()
control.pars$Htarg <- 0.1
control <- function(estimated.biomass, control.pars) {
  control.pars$Htarg
}

# Assume perfect implementation
implement <- function(TAC,...) {
  TAC
}

# Eval fxn - projects OM forward and implements mgmt procedure
evaluate <- function(pars.iter, biomass.iter,
  control.pars, data.years, proj.years,
  iterations, overshoot, ...) {
  # function arguments:
  # pars.iter & biomass.iter, the parameters & historical biomass trajectories of the operating model
  # control.pars, the specifications of the harvest control rule

  # set up some indexing values
  iyr <- length(data.years)+1
  pyr <- length(proj.years)
  yrs <- c(data.years, proj.years, max(proj.years)+1)
  # set up a data frame to store the results
  res <- data.frame()
  # loop over the iterations of the MSE, each iteration conducts a 20 year projection with annual generation of biomass
  # observations and appliations of the control rule.
  for(i in 1:iterations) {
    #i = 1

  #extract the parameters for this iteration
    K.i <- exp(pars.iter[i,1])
    r.i <- exp(pars.iter[i,2])
    sig.i <- exp(pars.iter[i,3])
    #set up vectors for time series of interest.
    biomass.i <- c(subset(biomass.iter, iter==i)$biomass, numeric(pyr))
    index.i <- c(index,numeric(pyr))
    catch.i <- c(harvest, numeric(pyr))
    TAC.i <- numeric(pyr)
    # loop over the projection period.
    for (y in iyr:(iyr+pyr-1)) {
      #generate the data for the most recent year
      index.i[y] <- observe(biomass.i[y] , sig.i)
      #calculate the TAC based on the harvest control rule
      # note that the control rule ONLY sees the index data, not the operating model biomass.
      TAC.i [y] <- control(index.i[y], control.pars) * index.i[y]
      #find the realized catch after implementation error
      catch.i[y] <- implement(TAC.i[y], overshoot)
      # update the true biomass of the operating model based on the output of the HCR
      biomass.i[y+1] <- schaefer(biomass.i[y],catch.i[y],K.i,r.i)
      #end projection year loop for iteration i
    }
    #store the results for this iteration
    res <- rbind(res, data.frame(year = yrs[-length(yrs)],
                                 value = index.i, type = "index", iter = i),
                 data.frame(year = yrs[-length(yrs)],
                            value = catch.i, type = "catch", iter=i),
                 data.frame(year = yrs, value = biomass.i,
                            type= "biomass", iter=i))
    #end loop over iterations
    }
  return(res)
# #end function evaluate()
}


```

```{r project_fix10percent}
project.fixed <- evaluate(pars.iter, biomass.iter, control.pars, data.years,
                          proj.years, niter, overshoot = NA)
tail(project.fixed)

projection.plot <- function(project.results) {
  #Fig2 <- ggplot(data = subset(project.results, type != "index"),
  # aes(x = year, y = value))
  project.results %>%
    filter(type %in% c("biomass","catch")) %>%
    group_by(type, year) %>%
    median_qi(value, .width = c(.5, .8, .95)) %>%
    ggplot() +
    geom_lineribbon(aes(x = year, y = value, ymin = .lower, ymax = .upper),
                    show.legend = FALSE) +
    scale_fill_brewer() +
    geom_line(aes(y=value,x=year),data = subset(project.results, type != "index" & iter==1 & year %in%
                                                  proj.years), lty=1,lwd=1,col=gray(0.7)) +
    geom_line(aes(y=value,x=year),data = subset(project.results, type != "index" & iter==2 & year %in%
                                                  proj.years), lty=1,lwd=1,col=gray(0.7)) +
    geom_line(aes(y=value,x=year),data = subset(project.results, type != "index" & iter==3 & year %in%
                                                  proj.years), lty=1,lwd=1,col=gray(0.7)) +
    #stat_summary(fun.data = "median_hilow", geom = "smooth", col="black",
    # fill = gray(0.5), lty = 2, aes=0.1) +
    # stat_summary(fun = median, fun.min = function(x)0, geom="line",
    # data = subset(project.results, type != "index" & year %in% data.years), lwd=1)
    facet_wrap(~type, scale = "free_y") +
    ylab("Million tonnes") +
    theme_bw()
}

projection.plot(project.fixed)


```

```{r MSE_altHCRs}
# Define a HCR that converts estimated biomassinto a harvest rate
control <- function(estimated.biomass, control.pars) {
  H1 <- control.pars$H1
  H2 <- control.pars$H2
  Bmax <- control.pars$Bmax
  B2 <- control.pars$B2
  B1 <- control.pars$B1
  harv <- ifelse(estimated.biomass >= B1, H1,
                 ifelse(estimated.biomass < B2, H2,
                        (H1-H2)/(B1-B2)*(estimated.biomass - B2) + H2))
  return(harv)
  #end function control
  }

# Define control params for HCR
control.pars <- list()
control.pars$H1 <- 0.05
control.pars$H2 <- 0
control.pars$Bmax <- max(index, na.rm =TRUE)
control.pars$B2 <- 0.2*control.pars$Bmax
control.pars$B1 <- 0.5*control.pars$Bmax

plot(c(0,control.pars$B2,control.pars$B1,control.pars$Bmax),
     c(control.pars$H2,control.pars$H2,control.pars$H1,control.pars$H1),
     type='l',axes=F,xlab="estimated biomass",ylab="exploitation rate",
     ylim=c(0,1.2*control.pars$H1))
axis(1,at=c(control.pars$B2,control.pars$B1),labels=c("B2","B1"))
axis(2,at=c(control.pars$H2,control.pars$H1),labels=c("H2","H1"))
box()

project.hcr <- evaluate(pars.iter, biomass.iter, control.pars,
                        data.years, proj.years, niter, overshoot = NA)

projection.plot(project.hcr)

```

```{r compare_HCRs}

# add potential to overshoot TAC
implement <- function(TAC, overshoot, ...) {
  TAC * (1 + overshoot)
}

control.pars <- list()
control.pars$H1 <- 0.05
control.pars$H2 <- 0
control.pars$Bmax <- max(index, na.rm =TRUE)
control.pars$B2 <- 0.2*control.pars$Bmax
control.pars$B1 <- 0.5*control.pars$Bmax

# Base Scenario (no TAC overshoot)
proj.hcr1.noerror <- evaluate(pars.iter, biomass.iter,
                              control.pars, data.years,
                              proj.years, niter,
                              overshoot = 0)

# 20% overshoot
proj.hcr1.error <- evaluate(pars.iter, biomass.iter,
                            control.pars, data.years,
                            proj.years, niter,
                            overshoot = 0.2)

control.pars$H1 <- 0.15

# Both scenarios, new target harvest rate
proj.hcr2.noerror <- evaluate(pars.iter, biomass.iter,
                              control.pars, data.years,
                              proj.years, niter,
                              overshoot = 0)
proj.hcr2.error <- evaluate(pars.iter, biomass.iter,
                            control.pars, data.years,
                            proj.years, niter,
                            overshoot = 0.2)


```

```{r diagnostics}
# Evaluate the different HCRs and create object with all results
MSE <- rbind(cbind(proj.hcr1.noerror, HCR="hcr1",
                   implement = "no overshoot"),
             cbind(proj.hcr1.error, HCR="hcr1",
                   implement = "20% overshoot"),
             cbind(proj.hcr2.noerror, HCR="hcr2",
                   implement = "no overshoot"),
             cbind(proj.hcr2.error, HCR="hcr2",
                   implement = "20% overshoot"))
head(MSE)

#Biomass and catch summed for all 4 options
Fig5 <- ggplot(data=subset(MSE, type !="index" &
                             year %in% proj.years),
               aes(x=HCR, y=value, ymin=0))
Fig5 + geom_boxplot(aes(fill=implement), width = 1) + facet_wrap(~type, scale="free_y") + ylab("Million Something")

#Yield based metrics (e.g. average annual catch)
#Stock Biomass metrics (e.g. distribution for B/BMSY, P(B>BLIM), etc.)
#Inter-annual stability of catch advice (e.g. how often the control rule closes the fishery)
aac2 <- with(MSE[MSE$year>max(data.years) & MSE$type=="catch",],
             aggregate(value,by=list(iter=iter,HCR=HCR,implement=implement),FUN=mean,na.rm=TRUE))
Fig6 <- ggplot(data=subset(aac2),
               aes(x=HCR, y=x, ymin=0))
Fig6 + geom_boxplot(aes(fill=implement), width = 1) + ylab("Million tonnes") + scale_fill_grey(start=0)

# num years fishery is open
not.closed <- with(MSE[MSE$year>max(data.years) & MSE$type=="catch",],
                   aggregate(value,by=list(iter=iter,HCR=HCR,implement=implement),FUN=num.above,threshold=0))
Fig8 <- ggplot(data=subset(not.closed),
               aes(x=HCR, y=x, ymin=0))
Fig8 + geom_boxplot(aes(fill=implement), width = 1) + ylab("Proportion of years TAC > 0") + scale_fill_grey(start = 0)

```












