---
title: "HighCatchExplore"
author: "Sophie Wulfing"
date: "2026-06-02"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(dplyr)
library(ggplot2)
library(tidyverse)

setwd('C:/Users/swulfing/Documents/GitHub/UMassD/YT_proj/Manuscripts/Draft2')
Proj_ests <- readRDS('Projections_data.rds')
om_ecov <- readRDS('C:/Users/swulfing/Documents/GitHub/UMassDartmouth-MSE/SW/om_ecov.RDS')


```

```{r quantiles, echo = FALSE}

SSB <- Proj_ests %>%
  filter(Year <= 2025) %>%
  group_by(EM_name, Year) %>%
  summarize(
    as_tibble_row(quantile(SSB))
  )
SSB <- pivot_longer(SSB, cols = c("0%", "25%", "50%", "75%", "100%"), names_to = 'Quantile', values_to = 'Value')
SSB$Quantile <- factor(SSB$Quantile, levels = c("0%", "25%", "50%", "75%", "100%"))
  
recruit <- Proj_ests %>%
  filter(Year <= 2025) %>%
  group_by(EM_name, Year) %>%
  summarize(
    as_tibble_row(quantile(OM_NAA_youngest))
  )

recruit <- pivot_longer(recruit, cols = c("0%", "25%", "50%", "75%", "100%"), names_to = 'Quantile', values_to = 'Value')
recruit$Quantile <- factor(recruit$Quantile, levels = c("0%", "25%", "50%", "75%", "100%"))
  
catch <- Proj_ests %>%
  filter(Year <= 2025) %>%
  group_by(EM_name, Year) %>%
  summarize(
    as_tibble_row(quantile(Pred_catch))
  )

catch <- pivot_longer(catch, cols = c("0%", "25%", "50%", "75%", "100%"), names_to = 'Quantile', values_to = 'Value')
catch$Quantile <- factor(catch$Quantile, levels = c("0%", "25%", "50%", "75%", "100%"))

fbar <- Proj_ests %>%
  filter(Year <= 2025) %>%
  group_by(EM_name, Year) %>%
  summarize(
    as_tibble_row(quantile(FBar))
  )

fbar <- pivot_longer(fbar, cols = c("0%", "25%", "50%", "75%", "100%"), names_to = 'Quantile', values_to = 'Value')
fbar$Quantile <- factor(fbar$Quantile, levels = c("0%", "25%", "50%", "75%", "100%"))

```

* I think this comes from a jump in recruitment around 2020 in the simulations in the 100th percentile, because immediately after that, we see a bump in SSB and a bump in catch/fbar following that.

* Catch is in kmt, recruitment is in individuals, I think SSB is also kmt but confirming that

* We don't see this high catch in the original OM that (see figure 10) we later see in the simulations. There's a little bump in recruitment toward the end followed by a little bump in fbar but not at all to the scale we see in some of the sims

* I compared this with the simulations I did with reduced recruitment sigma (from ~0.6 to 0.2):
    + SSB: lower in the reduced sigma simulations, the spike we see in the 100th quartile is more gradual in the reduced sigma on recruitment
    + Recruitment: lower in reduced sigma simulations, We don't see a spike at the end of the historical period
    + Catch: generally lower in reduced sigma simulations (with some exceptions)
    + Fbar: lower peak in reduced sigma simulations but still has one across projections

<!-- * According to the WHAM function reference, I don't really see anywhere where indices could have been weighted directly. There are places to specify not to use certain index values or you could alter the observation error on indices. -->

* Another solution could be to shorten historical period and see what that does (i.e. begin projecting at like 2015 instead of 2022)

```{=tex}
\newpage
```

```{r ssbQuantiles, fig.cap = 'SSB Quantiles', message = FALSE, warning = FALSE, echo = FALSE, out.height = '45%', fig.pos = 't'}

ggplot(SSB, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1) +
  theme_minimal()+ theme(legend.position = "none")

```

```{r ssb2Quantiles, fig.cap = 'SSB Quantiles rescaled without 100th percentile', message = FALSE, warning = FALSE, echo = FALSE, out.height = '45%', fig.pos = 'b'}

SSB2 <- SSB %>% filter(Quantile != '100%')
ggplot(SSB2, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1, scales = 'free_y') +
  theme_minimal()+ theme(legend.position = "none")
```

```{=tex}
\newpage
```

```{r recruitQuantiles, fig.cap = 'Recruitment Quantiles', echo = FALSE, message = FALSE, warning = FALSE, out.height = '45%'}

ggplot(recruit, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1) +
  theme_minimal()+ theme(legend.position = "none")

```

```{r recruit2Quantiles, fig.cap = 'Recruitment Quantiles rescaled without 100th percentile', echo=FALSE, message = FALSE, warning = FALSE, out.height = '45%'}

recruit2 <- recruit %>% filter(Quantile != '100%')
ggplot(recruit2, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1, scales = 'free_y') +
  theme_minimal()+ theme(legend.position = "none")
```

```{=tex}
\newpage
```

```{r catchQuantiles, fig.cap = 'Catch Quantiles', echo = FALSE, message = FALSE, warning = FALSE, out.height = '45%'}

ggplot(catch, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1) +
  theme_minimal()+ theme(legend.position = "none")
```

```{r catch2Quantiles, fig.cap = 'Catch Quantiles rescaled without 100 percentile', message = FALSE, warning = FALSE, echo = FALSE, out.height = '45%'}

catch2 <- catch %>% filter(Quantile != '100%')
ggplot(catch2, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1, scales = 'free_y') +
  theme_minimal()+ theme(legend.position = "none")
```

```{=tex}
\newpage
```

```{r fbarQuantiles, fig.cap = 'FBar Quantiles', echo = FALSE, message = FALSE, warning = FALSE, out.height = '45%'}

ggplot(fbar, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1) +
  theme_minimal()+ theme(legend.position = "none")

```

```{r fbar2Quantiles, fig.cap = 'FBar Quantiles rescaled without 100 percentile', echo = FALSE, message = FALSE, warning = FALSE, out.height = '45%'}

fbar2 <- fbar %>% filter(Quantile != '100%')
ggplot(fbar2, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1, scales = 'free_y') +
  theme_minimal()+ theme(legend.position = "none")
```

```{=tex}
\newpage
```

```{r fbarDiverge, fig.cap = 'This isnt that important Im just curious about the diverging trends in fbar based on the projections above', message = FALSE, warning = FALSE, echo = FALSE}


fbarDiv <- Proj_ests %>%
  filter(Year > 2025) %>%
  group_by(EM_name, Year) %>%
  summarize(
    as_tibble_row(quantile(FBar))
  )

fbarDiv <- pivot_longer(fbarDiv, cols = c("0%", "25%", "50%", "75%", "100%"), names_to = 'Quantile', values_to = 'Value')
fbarDiv$Quantile <- factor(fbarDiv$Quantile, levels = c("0%", "25%", "50%", "75%", "100%"))


ggplot(fbarDiv, aes(x = Year, y = Value, color = EM_name, group = EM_name)) +
  geom_line() +
  facet_wrap(~ Quantile, ncol = 1, scales = 'free_y') +
  theme_minimal()

```

```{=tex}
\newpage
```

```{r OGOM, fig.cap = 'Results from the original OM that we fitted with wham to use as base We dont really see the bump that we later see in the different simulations', message = FALSE, warning = FALSE, echo = FALSE}

library(patchwork)

OGOM <- data.frame(
  Years = om_ecov[["years"]],
  SSB = om_ecov[["rep"]][["SSB"]],
  catch = om_ecov[["rep"]][["pred_catch"]],
  recruit = om_ecov[["rep"]][["NAA"]][,,,1],
  fbar = om_ecov[["rep"]][["Fbar"]][,1]
)

p1 <- ggplot(OGOM, aes(x = Years, y = SSB)) + geom_line() + ylab('SSB')
p2 <- ggplot(OGOM, aes(x = Years, y = catch)) + geom_line() + ylab('Catch')
p3 <- ggplot(OGOM, aes(x = Years, y = recruit)) + geom_line() + ylab('Recruitment')
p4 <- ggplot(OGOM, aes(x = Years, y = fbar)) + geom_line() + ylab('fbar')

# Stack all 4 ggplots vertically
stacked_plot <- p1 / p2 / p3 / p4

# Display the stacked plot
stacked_plot

```










