---
title: "EcovContributionToRecruit"
author: "Sophie Wulfing"
date: "2026-06-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(dplyr)
library(ggplot2)
library(tidyverse)

setwd('C:/Users/swulfing/Documents/GitHub/UMassD/YT_proj/Manuscripts/Draft2')
Proj_ests <- readRDS('Projections_data.rds')

```

# The Calculation

log(R_t) = log(BH(SSB_t)) + beta * Ecov_t + epsilon_t
Then you can calculate:

Ecov contribution = beta * Ecov_t  (I assume you know what beta is in your OM and Ecov true state trajectory in your OM, but let me know if you can't find them)

and you could compare its variance to the total modeled variation in log recruitment:
Percent variation explained by Ecov = Var(beta * Ecov_t) / Var(log(R_t)) (so basically ecov_var/full_var)

One thing you should prob know is the variability of Rec is driven by three 'stochastic' component: (1) SSB (because you have B-H SRR I guess?), and SSB_t is driven by annual variability in M, F, and most importantly NAA_devs; (2) Ecov, which is a product of ecov series x beta that can increase/decrease R_t, now you can see if you have a high beta, this term becomes dominant and can drive R_t dramatically; (3) epsilon_t which is sigma_R you extracted (you should use realized one maybe["om"]][["rep"]][["NAA_devs"]][1,1,,1])

betas[i] <- mods[[i]][["Mod1"]][["om"]][["par"]][["Ecov_beta_R"]]
ecov_t <- data.frame(mods[[i]][["Mod1"]][["om"]][["rep"]][["Ecov_x"]][1:101,1])
SSB <- data.frame(mods[[i]][["Mod1"]][["om"]][["rep"]][["SSB"]])
Epsilon <- data.frame(mods[[i]][["Mod1"]][["om"]][["rep"]][["NAA_devs"]][1,1,,1])

```{r cars}

EcovEffects <- Proj_ests %>%
  mutate(Ecov_contribution_OM = Ecov_beta_R_OM * Temp_OM) %>%
  mutate(Ecov_contribution_EM = Ecov_beta_R_EM * Temp_EM) %>%
  mutate(logRecruit_OM = log(OM_NAA_youngest)) %>%
  mutate(logRecruit_EM = log(EM_NAA_youngest)) %>%
  group_by(EM_name, seed_no) %>%
  summarize(VarEcov_OM = var(Ecov_contribution_OM, na.rm = TRUE),
            VarEcov_EM = var(Ecov_contribution_EM, na.rm = TRUE),
            VarRec_OM = var(logRecruit_OM, na.rm = TRUE),
            VarRec_EM = var(logRecruit_EM, na.rm = TRUE)) %>%
  mutate(EcovEffect_OM = VarEcov_OM/VarRec_OM) %>%
  mutate(EcovEffect_EM = VarEcov_EM/VarRec_EM) %>%
  select(EM_name, EcovEffect_OM, EcovEffect_EM) # seend_no


EcovEffects_boxplot <- pivot_longer(EcovEffects, cols = c("EcovEffect_OM", "EcovEffect_EM"), names_to = 'Model', values_to = 'Value')

ggplot(EcovEffects_boxplot, aes(x = EM_name, y = Value, fill = Model)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45))


just_checking <- EcovEffects_boxplot %>% filter(Model == 'EcovEffect_OM')

ggplot(just_checking, aes(x = EM_name, y = Value, fill = Model)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45))

```
































