---
title: "Intro to MSE"
output: html_document
date: "2025-04-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggdist)
library(Hmisc)
library(mvtnorm)
# 
# library(palmerpenguins)
# library(furrr)
# library(MASS)
# library(janitor)
# library(skimr)
# library(bookdown)
# library(RcppRoll)

 
#SEE SLIDE 3 IN 01 - Intro - MSE for readings
```

```{r om}

data.years <- 1991:2013
harvest <- c(0.1,3,15,52,76,139,95,93,84,93,86,103,104,
92,46,67,59,30,54,59,47,33,44)
index <- c(NA,NA,NA,NA,NA,NA,NA,NA,935,NA,1057,NA,678,NA,
420,NA,554,NA,458,NA,474,NA,280)


plot(data.years,index, pch=19,xlab="Year",ylab="Million tonnes (B/C)",
ylim=c(0,1200))
lines(data.years,harvest,lty=2,lwd=2)

```

```{r OMDevel}
# LOGISTIC PRODUCTION FXN
schaefer <- function(B,C,K,r) {
  #function schaefer takes the current biomass, a catch,
  #and the model parameters to compute next year's biomass
  res <- B + B * r * (1 - B/K) - C
  return(max(0.001,res)) # we add a constraint to prevent negative biomass
}


# BIOMASS PRODUCTION FXN
dynamics <- function(pars,C,yrs) {
  # dynamics takes the model parameters, the time series of catch,
  # & the yrs to do the projection over
  # first extract the parameters from the pars vector (we estimate K in log-space)
  K <- exp(pars[1])
  r <- exp(pars[2])
  # find the total number of years
  nyr <- length(C) + 1
  # if the vector of years was not supplied we create
  # a default to stop the program crashing
  if (missing(yrs)) yrs <- 1:nyr
  #set up the biomass vector
  B <- numeric(nyr)
  #intialize biomass at carrying capacity
  B[1] <- K
  # project the model forward using the schaefer model
  for (y in 2:nyr) {
    B[y] <- schaefer(B[y-1],C[y-1],K,r)
  }
  #return the time series of biomass
  return(B[yrs])
  #end function dynamics
}

# function to calculate the negative log-likelihood
nll <- function(pars,C,U) { #this function takes the parameters, the catches, and the index data
  sigma <- exp(pars[3]) # additional parameter, the standard deviation of the observation error
  B <- dynamics(pars,C) #run the biomass dynamics for this set of parameters
  Uhat <- B #calculate the predicted biomass index - here we assume an unbiased absolute biomass estimate
  output <- -sum(dnorm(log(U),log(Uhat),sigma,log=TRUE),na.rm=TRUE) #calculate the negative log-likelihood
return(output)
#end function nll
}

# FXN TO PERFORM ASSESSMENT AND ESTIMATE PARAMS
assess <- function(catch,index,calc.vcov=FALSE,pars.init) {
  # assess takes catch and index data, initial values for the parameters,
  # and a flag saying whether to compute uncertainty estimates for the model parameters
  #fit model
  # optim runs the function nll() repeatedly with differnt values for the parameters,
  # to find the values that give the best fit to the index data
  res <- optim(pars.init,nll,C=catch,U=index,hessian=TRUE)
  # store the output from the model fit
  output <- list()
  output$pars <- res$par
  output$biomass <- dynamics(res$par,catch)
  output$convergence <- res$convergence
  output$nll <- res$value
  if (calc.vcov)
    output$vcov <- solve(res$hessian)
return(output)
#end function assess
}

# NOW RUN ASSESSMENT
ini.parms <- c(log(1200), log(0.1), log(0.3)) # Initial param vector for log(K), log(r), log(sigma)

# FIT TO DATA
redfish <- assess(harvest,index,calc.vcov=TRUE,ini.parms)
redfish


# EXTRACT MAX LIKELIHOOD AND PARAM ESTIMATES
biomass.mle <- redfish$biomass
# print(biomass.mle)
pars.mle <- redfish$pars
# print(exp(pars.mle))



```

```{r HW5OMDevel, eval = FALSE}
# LOGISTIC PRODUCTION FXN
schaefer <- function(B,C,K,r) {
  #function schaefer takes the current biomass, a catch,
  #and the model parameters to compute next year's biomass
  res <- B + B * r * (1 - B/K) - C
  res <- rnorm(1, mean(res), sd(harvest))
  return(max(0.001,res)) # we add a constraint to prevent negative biomass
}


# BIOMASS PRODUCTION FXN
dynamics <- function(pars,C,yrs) {
  # dynamics takes the model parameters, the time series of catch,
  # & the yrs to do the projection over
  # first extract the parameters from the pars vector (we estimate K in log-space)
  K <- exp(pars[1])
  r <- exp(pars[2])
  # find the total number of years
  nyr <- length(C) + 1
  # if the vector of years was not supplied we create
  # a default to stop the program crashing
  if (missing(yrs)) yrs <- 1:nyr
  #set up the biomass vector
  B <- numeric(nyr)
  #intialize biomass at carrying capacity
  B[1] <- K
  # project the model forward using the schaefer model
  for (y in 2:nyr) {
    B[y] <- schaefer(B[y-1],C[y-1],K,r)
  }
  #return the time series of biomass
  return(B[yrs])
  #end function dynamics
}

# function to calculate the negative log-likelihood
nll <- function(pars,C,U) { #this function takes the parameters, the catches, and the index data
  sigma <- exp(pars[3]) # additional parameter, the standard deviation of the observation error
  B <- dynamics(pars,C) #run the biomass dynamics for this set of parameters
  Uhat <- B #calculate the predicted biomass index - here we assume an unbiased absolute biomass estimate
  output <- -sum(dnorm(log(U),log(Uhat),sigma,log=TRUE),na.rm=TRUE) #calculate the negative log-likelihood
return(output)
#end function nll
}

# FXN TO PERFORM ASSESSMENT AND ESTIMATE PARAMS
assess <- function(catch,index,calc.vcov=FALSE,pars.init) {
  # assess takes catch and index data, initial values for the parameters,
  # and a flag saying whether to compute uncertainty estimates for the model parameters
  #fit model
  # optim runs the function nll() repeatedly with differnt values for the parameters,
  # to find the values that give the best fit to the index data
  res <- optim(pars.init,nll,C=catch,U=index,hessian=TRUE)
  # store the output from the model fit
  output <- list()
  output$pars <- res$par
  output$biomass <- dynamics(res$par,catch)
  output$convergence <- res$convergence
  output$nll <- res$value
  if (calc.vcov)
    output$vcov <- solve(res$hessian)
return(output)
#end function assess
}

# NOW RUN ASSESSMENT
ini.parms <- c(log(1200), log(0.1), log(0.3)) # Initial param vector for log(K), log(r), log(sigma)

# FIT TO DATA
redfish <- assess(harvest,index,calc.vcov=TRUE,ini.parms)
redfish


# EXTRACT MAX LIKELIHOOD AND PARAM ESTIMATES
biomass.mle <- redfish$biomass
# print(biomass.mle)
pars.mle <- redfish$pars
# print(exp(pars.mle))



```

```{r sampleparams}

# To obtain a set of plausible alternatives for the parameters of the operating model, we will use the statistical uncertainty from the estimation by sampling parameter sets from the estimated variance-covariance matrix.



#define the number of iterations for the MSE
niter <- 200
#set up a storage matrix for our alternative parameter sets
pars.iter <- matrix(NA,nrow = niter, ncol=3)
colnames(pars.iter) <- c("log_K","log_r","log_sigma")
# generate the sets of parameter values
for (i in 1:niter) {
  pars.iter[i,] <- mvtnorm::rmvnorm(1, mean = redfish$pars,
                                    sigma = redfish$vcov)
}
# Now generate replicate model outputs
biomass.iter <- data.frame()
for (i in 1:niter) {
  #here we calculate the biomass trajectory for each of the above sampled parameter vectors
  biomass.iter <- rbind(biomass.iter,
                        data.frame(year = seq(min(data.years),
                                              max(data.years)+1),
                                   biomass = dynamics(pars.iter[i,], harvest),
                                   iter = i))
}
biomass.iter

```

```{r plotBiomassTimeSeries}

biomass.iter %>%
  group_by(year) %>%
  median_qi(biomass, .width = c(.5, .8, .95)) %>%
  ggplot() +
  geom_lineribbon(aes(x = year, y = biomass, ymin = .lower, ymax = .upper),
                  show.legend = FALSE) +
  scale_fill_brewer() +
  # theme_bw() +
  geom_line(aes(y=harvest,x=year), data = tibble(harvest = harvest,
                                                 year = data.years),lty=2) +
  geom_point(aes(y=index, x=year), data = data.frame(index=index,
                                                     year = data.years)) +
  # geom_line(aes(y=biomass,x=year,group=iter,col=iter),data = subset(biomass.iter,iter%in%1:10)) +
  ylab("Estimated B and C (million tonnes)") +
  theme_bw() +
  guides(scale = "none")

```

```{r applyMSE}

proj.years <- 2014:2034

##### Data generation
observe <- function(biomass, sigma) {
  biomass * exp(rnorm(1, -0.5*(sigma**2), sigma))
}

```

```{r harvestControlRule}
# USING A FIXED TARGET EXPLOITATION RATE - calculates catch for next hear based on fixed percentage of most recent biomass estimate


control.pars <- list()
control.pars$Htarg <- 0.1
control <- function(estimated.biomass, control.pars) {
control.pars$Htarg
}

implement <- function(TAC,...) {
  TAC
}

evaluate <- function(pars.iter, biomass.iter,
                     control.pars, data.years, proj.years,
                     iterations, overshoot, ...) {
  # function arguments:
  # pars.iter & biomass.iter, the parameters & historical biomass trajectories of the operating model
  # control.pars, the specifications of the harvest control rule
  # set up some indexing values
  iyr <- length(data.years)+1
  pyr <- length(proj.years)
  yrs <- c(data.years, proj.years, max(proj.years)+1)
  # set up a data frame to store the results
  res <- data.frame()
  # loop over the iterations of the MSE, each iteration conducts a 20 year projection with annual generation of biomass
  # observations and appliations of the control rule.
  for(i in 1:iterations) { 
    #i = 1
    #extract the parameters for this iteration
    K.i <- exp(pars.iter[i,1])
    r.i <- exp(pars.iter[i,2])
    sig.i <- exp(pars.iter[i,3])
    #set up vectors for time series of interest.
    biomass.i <- c(subset(biomass.iter, iter==i)$biomass, numeric(pyr))
    index.i <- c(index,numeric(pyr))
    catch.i <- c(harvest, numeric(pyr))
    TAC.i <- numeric(pyr)
    # loop over the projection period.
    for (y in iyr:(iyr+pyr-1)) {
      #generate the data for the most recent year
      index.i[y] <- observe(biomass.i[y] , sig.i)
      #calculate the TAC based on the harvest control rule
      # note that the control rule ONLY sees the index data, not the operating model biomass. WHYYYYYYYYYYYYYYYYYYYYYYYYYYYYYY
      TAC.i [y] <- control(index.i[y], control.pars) * index.i[y]
      #find the realized catch after implementation error
      catch.i[y] <- implement(TAC.i[y], overshoot)
      # update the true biomass of the operating model based on the output of the HCR
      biomass.i[y+1] <- schaefer(biomass.i[y],catch.i[y],K.i,r.i)
      #end projection year loop for iteration i
      }
    #store the results for this iteration
    res <- rbind(res, data.frame(year = yrs[-length(yrs)],
                                 value = index.i, type = "index", iter = i),
                 data.frame(year = yrs[-length(yrs)],
                            value = catch.i, type = "catch", iter=i),
                 data.frame(year = yrs, value = biomass.i,
                            type= "biomass", iter=i))
    #end loop over iterations
    }
  res
  return(res)
  # #end function evaluate()
  }

```

```{r HW3harvestControlRule, eval = FALSE}
# HW3: Implement the HCR every 3 yrs rather than every 1.

control.pars <- list()
control.pars$Htarg <- 0.1
control <- function(estimated.biomass, control.pars) {
control.pars$Htarg
}

implement <- function(TAC,...) {
  TAC
}

evaluate <- function(pars.iter, biomass.iter,
                     control.pars, data.years, proj.years,
                     iterations, overshoot, ...) {
  # function arguments:
  # pars.iter & biomass.iter, the parameters & historical biomass trajectories of the operating model
  # control.pars, the specifications of the harvest control rule
  # set up some indexing values
  iyr <- length(data.years) + 1
  pyr <- length(proj.years)
  yrs <- c(data.years, proj.years, max(proj.years) + 1)
  # set up a data frame to store the results
  res <- data.frame()
  # loop over the iterations of the MSE, each iteration conducts a 20 year projection with annual generation of biomass
  # observations and appliations of the control rule.
  for(i in 1:iterations) { 
    #i = 1
    #extract the parameters for this iteration
    K.i <- exp(pars.iter[i,1])
    r.i <- exp(pars.iter[i,2])
    sig.i <- exp(pars.iter[i,3])
    #set up vectors for time series of interest.
    biomass.i <- c(subset(biomass.iter, iter==i)$biomass, numeric(pyr))
    index.i <- c(index,numeric(pyr))
    catch.i <- c(harvest, numeric(pyr))
    TAC.i <- numeric(pyr)
    # loop over the projection period.
    for (y in iyr:(iyr+pyr-1)) {
      #generate the data for the most recent year
      index.i[y] <- observe(biomass.i[y] , sig.i)
      #calculate the TAC based on the harvest control rule
      # note that the control rule ONLY sees the index data, not the operating model biomass. WHYYYYYYYYYYYYYYYYYYYYYYYYYYYYYY
      if(y %% 3 == 0){
        TAC.i [y] <- control(index.i[y], control.pars) * index.i[y]
        #find the realized catch after implementation error
        catch.i[y] <- implement(TAC.i[y], overshoot)
        # update the true biomass of the operating model based on the output of the HCR
        biomass.i[y+1] <- schaefer(biomass.i[y],catch.i[y],K.i,r.i)
        #end projection year loop for iteration i
      }
      else{ # THIS IS WHAT I CHANGED
        TAC.i [y] <- numeric(pyr)
        catch.i[y] <- c(harvest, numeric(pyr))
        # update the true biomass of the operating model based on the output of the HCR
        biomass.i[y+1] <- schaefer(biomass.i[y],catch.i[y],K.i,r.i)
      }
      }
    #store the results for this iteration
    res <- rbind(res, data.frame(year = yrs[-length(yrs)],
                                 value = index.i, type = "index", iter = i),
                 data.frame(year = yrs[-length(yrs)],
                            value = catch.i, type = "catch", iter=i),
                 data.frame(year = yrs, value = biomass.i,
                            type= "biomass", iter=i))
    #end loop over iterations
    }
  res
  return(res)
  # #end function evaluate()
  }

```

```{r HW4harvestControlRule, eval = FALSE}
# USING A FIXED TARGET EXPLOITATION RATE - calculates catch for next hear based on fixed percentage of most recent biomass estimate


control.pars <- list()
control.pars$Htarg <- 0.1
control <- function(estimated.biomass, control.pars) {
control.pars$Htarg
}

implement <- function(TAC,...) {
  rnorm(1, TAC, sd(harvest))
}

evaluate <- function(pars.iter, biomass.iter,
                     control.pars, data.years, proj.years,
                     iterations, overshoot, ...) {
  # function arguments:
  # pars.iter & biomass.iter, the parameters & historical biomass trajectories of the operating model
  # control.pars, the specifications of the harvest control rule
  # set up some indexing values
  iyr <- length(data.years)+1
  pyr <- length(proj.years)
  yrs <- c(data.years, proj.years, max(proj.years)+1)
  # set up a data frame to store the results
  res <- data.frame()
  # loop over the iterations of the MSE, each iteration conducts a 20 year projection with annual generation of biomass
  # observations and appliations of the control rule.
  for(i in 1:iterations) { 
    #i = 1
    #extract the parameters for this iteration
    K.i <- exp(pars.iter[i,1])
    r.i <- exp(pars.iter[i,2])
    sig.i <- exp(pars.iter[i,3])
    #set up vectors for time series of interest.
    biomass.i <- c(subset(biomass.iter, iter==i)$biomass, numeric(pyr))
    index.i <- c(index,numeric(pyr))
    catch.i <- c(harvest, numeric(pyr))
    TAC.i <- numeric(pyr)
    # loop over the projection period.
    for (y in iyr:(iyr+pyr-1)) {
      #generate the data for the most recent year
      index.i[y] <- observe(biomass.i[y] , sig.i)
      #calculate the TAC based on the harvest control rule
      # note that the control rule ONLY sees the index data, not the operating model biomass. WHYYYYYYYYYYYYYYYYYYYYYYYYYYYYYY
      TAC.i [y] <- control(index.i[y], control.pars) * index.i[y]
      #find the realized catch after implementation error
      catch.i[y] <- implement(TAC.i[y], overshoot)
      # update the true biomass of the operating model based on the output of the HCR
      biomass.i[y+1] <- schaefer(biomass.i[y],catch.i[y],K.i,r.i)
      #end projection year loop for iteration i
      }
    #store the results for this iteration
    res <- rbind(res, data.frame(year = yrs[-length(yrs)],
                                 value = index.i, type = "index", iter = i),
                 data.frame(year = yrs[-length(yrs)],
                            value = catch.i, type = "catch", iter=i),
                 data.frame(year = yrs, value = biomass.i,
                            type= "biomass", iter=i))
    #end loop over iterations
    }
  res
  return(res)
  # #end function evaluate()
  }

```

```{r project}

# Fixed 10% exploitation rate of est. biomass for all iterations and 20 yrs
project.fixed <- evaluate(pars.iter, biomass.iter, control.pars, data.years,
                          proj.years, niter, overshoot = NA)
tail(project.fixed)

# View trajectories of catch and om biomass from output
projection.plot <- function(project.results) {
  # Fig2 <- ggplot(data = subset(project.results, type != "index"),
  #  aes(x = year, y = value))
  project.results %>%
    filter(type %in% c("biomass","catch")) %>%
    group_by(type, year) %>%
    median_qi(value, .width = c(.5, .8, .95)) %>%
    ggplot() +
    geom_lineribbon(aes(x = year, y = value, ymin = .lower, ymax = .upper),
                    show.legend = FALSE) +
    scale_fill_brewer() +
    geom_line(aes(y=value,x=year),data = subset(project.results, type != "index" & iter==1 & year %in% proj.years), lty=1,lwd=1,col=gray(0.7)) +
    geom_line(aes(y=value,x=year),data = subset(project.results, type != "index" & iter==2 & year %in% proj.years), lty=1,lwd=1,col=gray(0.7)) +
    geom_line(aes(y=value,x=year),data = subset(project.results, type != "index" & iter==3 & year %in% proj.years), lty=1,lwd=1,col=gray(0.7)) +
    # stat_summary(fun.data = "median_hilow", geom = "smooth", col="black",
    #              fill = gray(0.5), lty = 2, aes=0.1) +
    # stat_summary(fun = median, fun.min = function(x)0, geom="line",
    #              data = subset(project.results, type != "index" & year %in% data.years), lwd=1) +
    facet_wrap(~type, scale = "free_y") +
    ylab("Million tonnes") +
    theme_bw()
}


projection.plot(project.fixed)

# THIS IS NOT COMING UP WITH MANAGEMENT SCENARIOS. WHYYY
```

```{r altHarvestControl}

control <- function(estimated.biomass, control.pars) {
  H1 <- control.pars$H1
  H2 <- control.pars$H2
  Bmax <- control.pars$Bmax
  B2 <- control.pars$B2
  B1 <- control.pars$B1
  harv <- ifelse(estimated.biomass >= B1, H1,
                 ifelse(estimated.biomass < B2, H2,
                        (H1-H2)/(B1-B2)*(estimated.biomass - B2) + H2))
  return(harv)
  #end function control
}

control.pars <- list()
control.pars$H1 <- 0.05
control.pars$H2 <- 0
control.pars$Bmax <- max(index, na.rm =TRUE)
control.pars$B2 <- 0.2*control.pars$Bmax
control.pars$B1 <- 0.5*control.pars$Bmax

# Plot HCR shape
plot(c(0,control.pars$B2,control.pars$B1,control.pars$Bmax),
     c(control.pars$H2,control.pars$H2,control.pars$H1,control.pars$H1),
     type='l',axes=F,xlab="estimated biomass",ylab="exploitation rate",
     ylim=c(0,1.2*control.pars$H1))
axis(1,at=c(control.pars$B2,control.pars$B1),labels=c("B2","B1"))
axis(2,at=c(control.pars$H2,control.pars$H1),labels=c("H2","H1"))
box()

project.hcr <- evaluate(pars.iter, biomass.iter, control.pars,
                        data.years, proj.years, niter)

# Plot HCR trajectories
projection.plot(project.hcr)

```

```{r compareHCRs}
# Just in case TAC is overshot
implement <- function(TAC, overshoot, ...) {
  TAC * (1 + overshoot)
}

# Compare different HCRs and account for if TAC is overshot
control.pars <- list()
control.pars$H1 <- 0.05
control.pars$H2 <- 0
control.pars$Bmax <- max(index, na.rm =TRUE)
control.pars$B2 <- 0.2*control.pars$Bmax
control.pars$B1 <- 0.5*control.pars$Bmax

# Base Scenario: no TAC overshoot
proj.hcr1.noerror <- evaluate(pars.iter, biomass.iter,
                              control.pars, data.years,
                              proj.years, niter, overshoot = 0)

# 20% Overshoot TAC
proj.hcr1.error <- evaluate(pars.iter, biomass.iter,
                            control.pars, data.years,
                            proj.years, niter,
                            overshoot = 0.2)

# Now change target harvest rate and run scenarios again with new target
control.pars$H1 <- 0.15

proj.hcr2.noerror <- evaluate(pars.iter, biomass.iter,
                              control.pars, data.years,
                              proj.years, niter,
                              overshoot = 0)
proj.hcr2.error <- evaluate(pars.iter, biomass.iter,
                            control.pars, data.years,
                            proj.years, niter,
                            overshoot = 0.2)

# Diagnostics
MSE <- rbind(cbind(proj.hcr1.noerror, HCR="hcr1",
                   implement = "no overshoot"),
             cbind(proj.hcr1.error, HCR="hcr1",
                   implement = "20% overshoot"),
             cbind(proj.hcr2.noerror, HCR="hcr2",
                   implement = "no overshoot"),
             cbind(proj.hcr2.error, HCR="hcr2",
                   implement = "20% overshoot"))
head(MSE)

# Summarize catch for all 4 options
Fig5 <- ggplot(data=subset(MSE, type !="index" &
                             year %in% proj.years),
               aes(x=HCR, y=value, ymin=0))
Fig5 + geom_boxplot(aes(fill=implement), width = 1) + facet_wrap(~type, scale="free_y") + ylab("Million")

#Yield based metrics (e.g. average annual catch)
#Stock Biomass metrics (e.g. distribution for B/BMSY, P(B>BLIM), etc.)
#Inter-annual stability of catch advice (e.g. how often the control rule closes the fishery)
aac2 <- with(MSE[MSE$year>max(data.years) & MSE$type=="catch",],
             aggregate(value,by=list(iter=iter,HCR=HCR,implement=implement),FUN=mean,na.rm=TRUE))
Fig6 <- ggplot(data=subset(aac2),
               aes(x=HCR, y=x, ymin=0))
Fig6 + geom_boxplot(aes(fill=implement), width = 1) + ylab("Million tonnes") + scale_fill_grey(start=0)

```

```{r performanceStats}
# years B > BLIM
# BLIM = 0.25*K (we specify BLIM for our performance as half BMSY)
blim <- 0.25*exp(pars.iter[,1])
num.above <- function(vec,threshold) {
  length(vec[vec>threshold])/length(vec)
}

MSE$blim <- blim[MSE$iter]
MSE$above.blim <- ifelse(MSE$value>MSE$blim,1,0)
above.blim <- with(MSE[MSE$year>max(data.years) & MSE$type=="biomass",],
                   aggregate(above.blim,by=list(iter=iter,HCR=HCR,implement=implement),FUN=sum,na.rm=TRUE))
above.blim$x <- above.blim$x/length(proj.years)

Fig7 <- ggplot(data=subset(above.blim),
               aes(x=HCR, y=x, ymin=0))
Fig7 + geom_boxplot(aes(fill=implement), width = 1) + ylab("Proportion of years above BLIM")

# num years fishery is open
not.closed <- with(MSE[MSE$year>max(data.years) & MSE$type=="catch",],
                   aggregate(value,by=list(iter=iter,HCR=HCR,implement=implement),FUN=num.above,threshold=0))

Fig8 <- ggplot(data=subset(not.closed),
               aes(x=HCR, y=x, ymin=0))
Fig8 + geom_boxplot(aes(fill=implement), width = 1) + ylab("Proportion of years TAC > 0")
```

# Next Steps - DO THESE AFTER COMPLETING THE WHOLE COURSE

Your turn to add features!

Suggestions:

1. Produce a trade-off plot (hint: perhaps think about some alternative performance statistics that integraten across iterations)

2. Add a model-based control rule by performing a stock assessment (e.g. production model) each year in the projection period. Then use the catch associated with the estimated FMSY as the TAC. Be careful not to give the assessment model the true parameter values from the operating model. - ALREADY IN MODEL

3. Implement the HCR every 3 yrs rather than every 1. - DONE

4. Add a more complicated implementation function (say based on price?) - DONE

5. Add environmental variability (process error) into the population dynamics - DONE?

